Placing the spotted T Tauri star LkCa 4 on an HR diagram

### 2016-08-10 document/ms.tex
Final round of edits to the draft before we set Version 0.5
Comments from GJH on the version 0.5 draft.

\begin{abstract}
The weak-lined T-Tauri star LkCa 4 has no evidence for IR excess, accretion, or nearby companions, and shows exceptionally large photometric modulation attributable to star spots, making it an ideal testbed for measuring starspot properties.  Infrared spectral features in high resolution data indicate that 70-80\% of the surface is covered in 2700$-$3000 K effective temperature regions at the epoch of the observation.  The photometric phase curve is scaled to this filling factor to construct the fill-factor-as-a-function-of-phase.  The SED, color variability, and low-resolution optical/IR spectroscopy are registered to filling factor at their epochs, confirming the large filling factor and temperature contrast of the starspots through time.  The revised effective temperature and luminosity make LkCa4 appear much lower mass and much younger when placed on a pre-main sequence HR diagram, compared to previous estimates based only on optical spectroscopy and not taking into account the large coverage fraction of cool spots.
\end{abstract}

@@

Magnetic activity is a likely source of significant uncertainty in both the models and observations of young low-mass stars. Convection at these young ages generate strong magnetic fields, as measured in Zeeman broadening and polarimetry \citep[e.g.][]{johnskrull07,donati09} and as seen in starspots \citep[e.g.][]{stauffer03,grankin08}, which cause significant jitter in radial velocity surveys for young planets \citep[e.g.][]{donati14, robertson14}.  Evolutionary models are just now starting to implement new prescriptions for convection with magnetic fields (\citet{somers15}, \citet{feiden16}; see also \citet{baraffe15} for an updated treatment of convection without introducing magnetic fields).  Stellar evolution models including the effect of starspots can make a coeval 10 Myr population exhibit apparent age spreads of 3$-$10 Myr, with derived masses biased towards lower masses \citep{somers15}.

@@

Starspots have recently been observed with exceptional photometric precision from monitoring surveys targeting exoplanet transits \citep{harrison11,davenport15}.  Contemporaneous or near-contemporaneous panchromatic photometric monitoring \citep{bouvier95,grankin07} can break the degeneracy between starspot areal filling factor and starspot temperature. Photometric measurements are limited to detecting large longitudinally asymmetric starspots.  Transiting planets or eclipsing binaries passing in front of starspots \citep{desert11} can break geometric degeneracies, providing estimates for the characteristic sizes and lifetimes of starspots or starspot groups.  Zeeman Doppler Imaging \citep[ZDI]{donati14} suffers from the same longitudinal asymmetry limitation as photometry, but can yield information on the latitudinal distribution of starspots.  All of these techniques would underestimate the starspot areal coverage of pole-on stars, stars with longitudinally symmetric bands of starspots, or stars with an isotropic distribution of small starspots.

Spectroscopic strategies have the power to measure both the starspot areal coverage and temperature in any geometric configuration.  At low resolution, SED modeling \citep{wolk96} has detected evidence for sunspots in weak-lined T-Tauri stars (WTTS).  At high resolution, line depth ratio analysis has been demonstrated mostly on evolved stars in relatively small bandwidths, namely the TiO bands \citep{neff95,oneal96,oneal98,oneal04}, or the OH 1.563 $\mu$m feature \citep{oneal01}.

Past spectroscopic detections of spots on young stars have focused on TW Hya, an active accretor, and DQ Tau, a close binary in which both components are accreting \citep{debes13,bary14}.  The WTTS \name \citep{herbig86,strom89a,downes88,strom89b} is an ideal exemplar for a spotted pre-MS star because it does not have any veiling \citep{hartigan95} or mid-IR or mm excess \citep{skrutskie90,andrews05,mccabe06,furlan06,luhman10,furlan11,harris12,marichalar12,howard13,buckle15}, is not actively accreting \citep{hartigan95,kenyon98,edwards06,cauley12,yang12,ardila13,castro13,loyd14,martinez15}, has no evidence for a close companion \citep{karr10,kraus11,daemgen15}\footnote{The status for wide companions is less clear; see \citet{stauffer91,itoh08,kraus09,kraus11,herczeg14}}, demonstrates large amplitude sinusoidal photometric variability indicative of large spots \citep{grankin08,xiao12}, has recently been examined with ZDI \citep{donati14}, and has large ($\Delta \teff \sim$500 K) disagreement about its effective temperature \citep{herczeg14}.  The spectrum of \name should be devoid of complicating factors like near-IR excess veiling and accretion excess.

@@

\subsection{IGRINS Spectroscopy}\label{sec:igrins} 
We acquired observations with the Immersion Grating Infrared Spectrograph, IGRINS \citep{gully12,park14} on the Harlan J. Smith Telescope at McDonald Observatory on 2015-11-18 $09^h$ UTC.  IGRINS is a high resolution near-infrared echelle spectrograph providing simultaneous $R\simeq45,000$ spectra over 1.48-2.48 \um.  The spectrograph has two arms with 28 orders in $H-$band and 25 orders in $K-$band.  The data were reduced with the Pipeline Package\footnote{\url{https://github.com/igrins/plp}}.  Telluric correction was performed by dividing the spectrum of \name by a spectrum of HR 1237, an A0V star observed immediately before \name, both at airmass 1.1.  The broad hydrogen lines in the A0V star produced broad flux excesses in the spectrum of LkCa 4.  No effort was made to remove these flux excesses, which affect several spectral orders.

@@

As noted previously, line depth ratio analysis has traditionally been limited to isolated portions of spectrum that possess spectral lines attributable only to starspots and spectral lines attributable only to ambient photosphere \citep[\emph{e.g.}][]{neff95, oneal01}.  The apparent veiling of the lines and their respective temperature dependences can be combined to solve for the starspot and ambient photosphere temperatures and relative areal coverages.  The line depth ratio method suffers from the need to identify portions of the spectrum that possess such strong lines, and also suffers from the need to assemble large atlases of observed spectral templates to which the spotted star spectrum can be compared.  In this section, we introduce a generalization of line depth ratio analysis which employs pixel-by-pixel modeling of spectral orders.  By leveraging all the channels present in a spectrum, this strategy has the power to constrain starspot properties at relatively low fill-factors and identify weak lines that originate from the starspot.

@@

\citet[hereafter \iancze]{czekala15} developed a modular framework\footnote{The open source codebase and its full revision history is available at \url{https://github.com/iancze/Starfish}.  The experimental fork discussed in this paper is at \url{https://github.com/gully/Starfish}} to infer stellar properties from high resolution spectra.  The \iancze\ technique forward models ``emulated'' synthetic spectra from pre-computed model grids, taking into account the discretization error attributable to the coarsely sampled stellar intrinsic parameters.  We employed the \PHOENIX\ grid of pre-computed synthetic stellar spectra that span a wide range of wavelengths at high spectral resolution with sampling of 100 K in $\teff$ in our range of interest \citep{husser13}.  The modular framework was altered to accommodate starspot measurements in two ways. First, the single photospheric component was updated to include a starspot spectrum. Second, the MCMC sampling strategy was altered to accommodate the additional free parameters added in the starspot model.

@@

Note that $f_{\Omega}$ represents an instantaneous, observational fill factor seen on one projected hemisphere, not $f_{\rm spot}$, the ``ratio of the spotted surface to the total surface areas'' \citep{somers15}.  In the limit of homogenously distributed spots, $f_{\Omega}$ tends to $f_{\rm spot}$, but in general $f_{\rm spot}$ is not a direct observable since some circumpolar regions of inclined stars will always face away from Earth; $f_{\Omega}$ can vary cyclically through rotational modulation, whereas $f_{\rm spot}$ changes secularly through starspot evolution that is not yet understood.

The starspot model includes two more free parameters than the standard \iancze\ model, namely $\teffb$ and $f_{\Omega}$.  The addition of these two parameters makes the MCMC sampling much more correlated than it was before because the relative contribution of the starspot and ambient photosphere are nearly degenerate over small changes in effective temperature.  So the second important change to the spectral inference framework involves technical aspects of switching from sampling the nuisance and stellar parameters separately in a blocked Gibbs framework to ensemble sampling using \texttt{emcee} \citep{foreman13}.  The affine-invariant \texttt{emcee} ensemble sampler is more resilient to correlations among stellar temperatures and fill factor than the Metropolis-Hastings sampler used in Gibbs sampling.  The practical effect of this switch is that all 14 stellar and nuisance parameters are fit simultaneously in a single spectral order.  So stellar parameter estimates $\vt_{o}$ are unique for each spectral order $o$, whereas the \iancze\ strategy had the power to provide a single set of stellar parameters $\vt$ that was based on all $N_{\rm ord}$ spectral orders.  The $N_{\rm ord}$ sets of inferences on $\vt_{o}$ are combined by weighted averaging point estimates of results from reliable orders, which offers resilience to flagrant calibration artifacts or conspicuous model mismatches.  This process can be thought of as a coarse approximation to the much more sophisticated spectral line outlier rejection described in \iancze.

@@

We find a larger dispersion in the point estimates for the stellar parameters derived from the IGRINS data than those derived in optical.  The most conspicuous trend is in the derived effective temperature as a function of wavelength shown in Figure \ref{fig:SingleTeffvsOrder}.  The effective temperature peaks at values of $\sim4200$K in the short wavelength end of $H-$band and saturates at $<3500$ K at the long wavelength end of $H-$band.  The $K-$band shows even lower derived effective temperatures of $\sim3300$ K, a $Delta \teff \sim 700$ K spread from optical.  No single temperature can describe all the spectral lines present in the high resolution optical and IR spectra.  This discrepancy is circumstantial evidence for the detection of spectral features attributable to starspots.

@@

In summary, the visible portion of the spectrum will be dominated by the warm patches, but the contrast with the starspots will be higher.  The infrared spectrum will have a lower overall spot contrast, but will have more net starspot flux than optical.  This observation has motivated precision radial velocity (RV) planet searches in the infrared, where the modulation of RV signals is lessened \citep{prato08}.

@@

Full-spectrum fitting was performed for 48 of the 54 available IGRINS spectral orders as described in Section \ref{sec:methods} and Appendix \ref{methods-details}.  The MCMC samples had 5000 steps with 40 walkers.  The stellar parameter ranges were $\logg \in [3.5, 4.0]$ and $\Z \in [-0.5, 0.5]$; the effective temperature range was $\teffa, \teffb \in [2700, 4500]$.  $H-$band fits had normal distribution priors in place: solar metallicity to $\pm0.05$ dex, $logg=3.8\pm0.1$, and $\vsini=29\pm5$.  The MCMC chains all appeared to converge after about 1500 steps.

Marginalized distributions were computed for all 14 stellar and nuisance parameters by selecting the final 200 $\times$ 40 walkers = 8000 samples, and point estimates were obtained by computing the median, $5^{th}$ and $95^{th}$ percentiles of the marginal distributions.  The fit quality was first assessed by examining the consistency of the point estimates across the spectral orders.  The distribution of $v_z$ and $\vsini$ were informative for finding extremely poor fits.  The range for all 48 orders was $v_z = 12.4 \pm 2.6$ and $\vsini = 28.8 \pm 2.0$.  Two orders with poor telluric correction ($o=91$ and $94$) had exceptionally large uncertainties on their values for $v_z$ and $\vsini$.

The posteriors on $\teffa$ and $\teffb$ show several trends- see Figure \ref{fig:TwoTempResults} and Appendix Table \ref{tbl_order_results} for the results of effective temperatures for all IGRINS orders.

Overplotting models with the observed spectra yielded insights on why some orders failed, and which orders can discriminate between starspots and ambient photospheric lines the best.  Some orders indicate that their spectrum is dominated by spectral line outliers--deep metal lines that cannot be described by a coarse metallicity estimate alone.  These spectral line outlier orders also tend to bias estimates of effective temperature.  These metal lines can also be biased by Zeeman broadening, which would explain the heightened effect size with wavelength.

@@

Figures \ref{fig:Hband3x7} and \ref{fig:Kband3x7} in the Appendix show 42 of the $H-$ and $K-$band spectra on a log scale with a single random composite model spectrum overplotted.  The red and blue lines show the hot photosphere and cool spot spectra respectively.  We inspected the quality of the fits, generally finding good agreement with most spectral orders.  Many orders demonstrated spectral line outliers.  Most of the gross spectral features are captured by the composite model.  There are many examples of spectral lines that are present only in the hot photosphere model, or only in the cool photosphere model.  For example, the OH 1.563 $\mu$m line noted by \citet{oneal01} shows a clear pattern of 3 lines in the data, with the central line exceeding the depth of the adjacent two lines.

Best fit values for the stellar parameters summarized are listed in Table \ref{tbl_adopted_props}.  Remarkably, the filling factor of starspots exceeds 50\% with a best fit value of $c=75\pm 10 \% $.

@@

The two panels in Figure \ref{fig:lores_twopanel} show photospheric spectra scaled separately to the observed red-optical and near-IR spectra.  The single temperature spectra are unable to reproduce the observed emission.  The $3900$ K component fits the TiO bands reasonably well at $<7400$ \AA\ but fails to reproduce molecular bands at longer wavelengths.  The $3500$ K component hotter component suffers from the opposite problem, yielding molecular bands at short wavelengths that are too deep.  On the best-fit photosphere from \S 3 fits both the optical and the near-IR bands well.

@@

The model photometry is obtained from the main sequence colors compiled by \citet{kenyon95}.  Colors from dwarf stars should have minimal contribution from spots; measurements of pre-main sequence colors will include some unknown contribution of spots and cannot be used for this analysis.  Synthetic spectra of $3000-4000$ K stars predict $B-V$ colors that are much bluer than observed and that become bluer at cooler temperatures, the opposite of what is observed.  As a consequence, synthetic photometry also cannot be used here.

@@

The rotational modulation leads to much smaller brightness changes at near-IR wavelengths than at optical wavelengths (see Table~\ref{tab:photrange}).  The smaller amplitude of near-IR brightness is also seen in a few spotted stars in the optical/IR monitoring of NGC 2264 \citep{cody14}, although those stars have a much smaller amplitude than LkCa 4 in $V$ band.  Figure~\ref{fig:colors} demonstrates that the $B-V$ color from \citet{grankin08} does not depend on $V$, while the $V-R_J$ becomes slightly redder when $V$ is smaller.  The predicted colors from rotational modulation do not match the observed $V-R_J$ colors or the correlation with $V$, either because of changes in spot properties since the \citet{grankin08} data were obtained or because a two-temperature model for the photosphere is overly simplistic.

@@

The TiO band depths in these spectra are measured from spectral indices listed in Table~\ref{tab:tio}.  The TiO indices are also combined into a single index by  calculating how many standard deviations each point is from the median value of the index, and then averaging the standard deviation.

@@

Scaling our composite spectrum for LkCa 4 to the observed $K$-band photometry yields $R=2.2$ $R_\odot$ and $\log L/L_\odot=0.17$, assuming a distance of 131 pc based the parallax distance of the nearest Taurus member \citep{torres12}.   The spectrum of LkCa 4 is well reproduced by a cool component of $3000$ K covering 75\% of the surface and a $4200$ K component covering 25\% of the surface.  The effective temperature of these two components is 3611 K, as calculated from the average surface flux at the median $V$ band magnitude.  The apparent effective temperature in any single epoch would depend on the relative fractions of the hot and cool spot.  Decreasing the cool spot size to 60\% would increase the effective temperature to 3670 K, while increasing the spot size to 85\% would decrease the effective temperature to 3370 K.  {\bf GJH note:  holy crap!!!}.  Since the Hayashi tracks are nearly vertical at this temperature and age, the temperature uncertainty leads directly to an uncertainty in inferred mass.

@@

@@

>>>>

\begin{abstract}
Starspots temporarily increase the opacity in stellar photospheres, impacting both their instantaneous stellar flux and the duration and efficiency of their convective energy transport during pre-main sequence stellar evolution.  The observational manifestation of local magnetic fields, starspots on young stars can cover a large fraction of the stellar surface, causing the incorrect assessment of both the effective temperature and the luminosity and therefore the placement of sources on HR diagrams.  The singleton weak-lined T-Tauri star LkCa 4 shows exceptionally large photometric modulation attributable to star spots, making it an ideal testbed for measuring starspot properties.  Infrared spectral features in high resolution IGRINS spectra indicate that 70-80\% of the surface is covered in $\teff \sim 2700-3000$ K patches of cool photosphere at the epoch of the observation.  The rest of the $20-30\%$ of the stellar surface can be thought of as either ambient photosphere, or a hot spot of $\teff\sim4200$ K.  The photometric phase curve is scaled to this filling factor to construct the fill-factor-as-a-function-of-phase.  The SED, color variability, and low-resolution optical/IR spectroscopy are registered to the filling factors at their epochs, strengthening the large filling factor and temperature contrast of the starspots through time.  The revised effective temperature and luminosity make LkCa4 appear much lower mass and much younger when placed on a pre-main sequence HR diagram, compared to previous estimates based only on optical spectroscopy and not taking into account the large coverage fraction of cool spots.  LkCa 4 represents the extreme end of starspot coverage, suggesting that many heavily spotted young stars have been misplaced on the observational HR diagram, leading to incorrect ages and masses and contributing, in part, to the apparent age spread observed towards young clusters.
\end{abstract}

@@

Magnetic activity is a likely source of significant uncertainty in both the models and observations of young low-mass stars. Convection at these young ages generates strong magnetic fields, as measured in Zeeman broadening and polarimetry \citep[e.g.][]{johnskrull07,donati09} and as seen in starspots \citep[e.g.][]{stauffer03,grankin08}.  Evolutionary models are just now starting to implement new prescriptions for convection with magnetic fields (\citet{somers15}, \citet{feiden16}; see also \citet{baraffe15} for an updated treatment of convection without introducing magnetic fields).  Stellar evolution models including the effect of starspots can make a coeval 10 Myr population exhibit apparent age spreads of 3$-$10 Myr, with derived masses biased towards lower masses \citep{somers15}.

@@

Starspots have recently been observed with exceptional photometric precision from monitoring surveys targeting exoplanet transits \citep[\emph{e.g}][]{harrison11,cody14,davenport15}.  Contemporaneous or near-contemporaneous panchromatic photometric monitoring \citep{bouvier95,grankin07} can break the degeneracy between starspot areal filling factor and starspot temperature. Photometric measurements are limited to detecting large longitudinally asymmetric starspots.  Transiting planets or eclipsing binaries passing in front of starspots \citep{desert11} can break geometric degeneracies, providing estimates for the characteristic sizes and lifetimes of starspots or starspot groups.  Zeeman Doppler Imaging \citep[ZDI]{donati14} suffers from the same longitudinal asymmetry limitation as photometry, but can yield information on the latitudinal distribution of starspots.  All of these techniques would underestimate the starspot areal coverage of pole-on stars, stars with longitudinally symmetric bands of starspots, or stars with an isotropic distribution of small starspots.

Spectroscopic strategies have the power to measure both the starspot areal coverage and temperature in any geometric configuration.  At low resolution, SED modeling \citep{wolk96} has detected evidence for sunspots in weak-lined T-Tauri stars (WTTS).  At high resolution, line depth ratio analysis has been demonstrated mostly on evolved stars in relatively small bandwidths, namely the TiO bands \citep{neff95,oneal96,oneal98,oneal04}, or the OH 1.563 $\mu$m feature \citep{oneal01}.  Past spectroscopic detections of spots on young stars have focused on TW Hya, an active accretor, and DQ Tau, a close binary in which both components are accreting \citep{debes13,bary14}.

The WTTS \name \citep{herbig86,strom89a,downes88,strom89b} is an ideal exemplar for a spotted pre-MS star because it does not have any veiling \citep[\emph{e.g.}][]{hartigan95} or mid-IR or mm excess \citep{skrutskie90,andrews05,mccabe06,furlan06,luhman10,furlan11,harris12,marichalar12,howard13,buckle15}, is not actively accreting \citep[\emph{e.g.}]{kenyon98,edwards06,cauley12,yang12}, has no evidence for a close companion \citep{karr10,kraus11,nguyen12,daemgen15}\footnote{The status for wide companions is less clear; see \citet{stauffer91,itoh08,kraus09,kraus11,herczeg14}}, and demonstrates large amplitude sinusoidal photometric variability indicative of large spots \citep{grankin08,xiao12}.    The spectrum of \name should be devoid of complicating factors like near-IR excess veiling and accretion excess.

LkCa 4 has recently been examined with ZDI \citep{donati14}.  Unanswered questions about the large disagreement ($\Delta \teff \sim$500 K) in LkCa 4's effective temperature \citep{herczeg14} remain.

@@

\subsection{IGRINS Spectroscopy}\label{sec:igrins} 
We acquired observations with the Immersion Grating Infrared Spectrograph, IGRINS \citep{park14} on the Harlan J. Smith Telescope at McDonald Observatory on 2015-11-18 $09^h$ UTC.  IGRINS is a high resolution near-infrared echelle spectrograph providing simultaneous $R\simeq45,000$ spectra over 1.48-2.48 \um.  The spectrograph has two arms with 28 orders in $H-$band and 25 orders in $K-$band \citep{gully12}.  The data were reduced with the Pipeline Package\footnote{\url{https://github.com/igrins/plp}}.  Telluric correction was performed by dividing the spectrum of \name by a spectrum of HR 1237, an A0V star observed immediately before \name, both at airmass 1.1.  The broad hydrogen lines in the A0V star produced broad flux excesses in the spectrum of LkCa 4.  No effort was made to remove these flux excesses, which affect several spectral orders.

@@

Starspots possess a spectrum distinct from the ambient photosphere.  Spatially resolved spectra of starspots have not yet been observed for stars beyond the sun, though interferometric techniques \todo{cite} and planetary transits \todo{cite} are promising.  Approaches like ZDI and line-bisector analysis \todo{cite} target the modulation of stellar photospheric lines as starspots enter and exit the observable stellar disk.  In contrast, the technique described in this Section detects and measures the spectral lines originating from starspots photospheric emission.

Starspot line depth ratio analysis has traditionally been limited to isolated portions of spectrum that possess spectral lines attributable only to starspots and spectral lines attributable only to ambient photosphere \citep[\emph{e.g.}][]{neff95, oneal01}.  The apparent veiling of the lines and their respective temperature dependences can be combined to solve for the starspot and ambient photosphere temperatures and relative areal coverages.  The line depth ratio method suffers from the need to identify portions of the spectrum that possess such strong lines, and also suffers from the need to assemble large atlases of observed spectral templates to which the spotted star spectrum can be compared.  In this section, we introduce a generalization of line depth ratio analysis which employs pixel-by-pixel modeling of spectral orders.  By leveraging all the channels present in a spectrum, this strategy has the power to constrain starspot properties at relatively low fill-factors and identify weak lines that originate from the starspot.

@@

\citet[hereafter \iancze]{czekala15} developed a modular framework\footnote{The open source codebase and its full revision history is available at \url{https://github.com/iancze/Starfish}.  The experimental fork discussed in this paper is at \url{https://github.com/gully/Starfish}} to infer stellar properties from high resolution spectra.  The \iancze\ technique forward models ``emulated'' synthetic spectra from pre-computed model grids, taking into account the discretization error attributable to the coarsely sampled stellar intrinsic parameters.  We employed the \PHOENIX\ grid of pre-computed synthetic stellar spectra, which span a wide range of wavelengths at high spectral resolution with sampling of 100 K in $\teff$ in our range of interest \citep{husser13}.  The modular framework was altered to accommodate starspot measurements in two ways. First, the single photospheric component was updated to include a starspot spectrum. Second, the MCMC sampling strategy was altered to accommodate the additional free parameters added in the starspot model.

@@

The term $f_{\Omega}$ represents an instantaneous, observational fill factor seen on one projected hemisphere, not $f_{\rm spot}$, the ``ratio of the spotted surface to the total surface areas'' \citep{somers15}.  In the limit of homogenously distributed spots, $f_{\Omega}$ tends to $f_{\rm spot}$, but in general $f_{\rm spot}$ is not a direct observable since some circumpolar regions of inclined stars will always face away from Earth; $f_{\Omega}$ can vary cyclically through rotational modulation, whereas $f_{\rm spot}$ changes secularly through starspot evolution that is not yet understood.

The starspot model includes two more free parameters than the standard \iancze\ model, namely $\teffb$ and $f_{\Omega}$.  The addition of these two parameters makes the MCMC sampling much more correlated than it was before because the relative contribution of the starspot and ambient photosphere are nearly degenerate over small changes in effective temperature.  This challenge motivated the second important change to the spectral inference framework, which involves technical aspects of switching from sampling the nuisance and stellar parameters separately in a blocked Gibbs framework to ensemble sampling using \texttt{emcee} \citep{foreman13}.  The affine-invariant \texttt{emcee} ensemble sampler is more resilient to correlations among stellar temperatures and fill factor than the Metropolis-Hastings sampler used in Gibbs sampling.  The practical effect of this switch is that all 14 stellar and nuisance parameters are fit simultaneously in a single spectral order.  Stellar parameter estimates $\vt_{o}$ are unique for each spectral order $o$, whereas the \iancze\ strategy had the power to provide a single set of stellar parameters $\vt$ that was based on all $N_{\rm ord}$ spectral orders.  The $N_{\rm ord}$ sets of inferences on $\vt_{o}$ are combined by weighted averaging point estimates of results from reliable orders, which offers resilience to flagrant calibration artifacts or conspicuous model mismatches.  This process can be thought of as a coarse approximation to the much more sophisticated spectral line outlier rejection described in \iancze.

@@

We find a larger dispersion in the point estimates for the stellar parameters derived from the IGRINS data than those derived in optical.  The most conspicuous trend is in the derived effective temperature as a function of wavelength shown in Figure \ref{fig:SingleTeffvsOrder}.  The effective temperature peaks at values of $\sim4200$K in the short wavelength end of $H-$band and saturates at $<3500$ K at the long wavelength end of $H-$band.  The $K-$band shows even lower derived effective temperatures of $\sim3300$ K, or 700 K cooloer than estimated from optical.  No single temperature can describe all the spectral lines present in the high resolution optical and IR spectra.  This discrepancy is circumstantial evidence for the detection of spectral features attributable to starspots.

@@

In summary, the visible portion of the spectrum will be dominated by the warm patches, but the contrast with the starspots will be higher.  The infrared spectrum will have a lower overall spot contrast, but will have more net starspot flux than optical.

@@

Full-spectrum fitting was performed for 48 of the 54 available IGRINS spectral orders as described in Section \ref{sec:methods} and Appendix \ref{methods-details}.  The MCMC samples had 5000 steps with 40 walkers.  The stellar parameter ranges were $\logg \in [3.5, 4.0]$ and $\Z \in [-0.5, 0.5]$; the effective temperature range was $\teffa, \teffb \in [2700, 4500]$.  $H-$band fits had normal distribution priors in place: solar metallicity to $\pm0.05$ dex, $logg=3.8\pm0.1$, and $\vsini=29\pm5$ km/s.  The MCMC chains all appeared to converge after about 1500 steps.

Marginalized distributions were computed for all 14 stellar and nuisance parameters by selecting the final 200 $\times$ 40 walkers = 8000 samples, and point estimates were obtained by computing the median, $5^{th}$ and $95^{th}$ percentiles of the marginal distributions.  The fit quality was first assessed by examining the consistency of the point estimates across the spectral orders.  The distribution of $v_z$ and $\vsini$ were informative for finding extremely poor fits.  The range for all 48 orders was $v_z = 12.4 \pm 2.6$ km/s and $\vsini = 28.8 \pm 2.0$ km/s.  Two orders with poor telluric correction ($o=91$ and $94$) had exceptionally large uncertainties on their values for $v_z$ and $\vsini$.

The posteriors on $\teffa$ and $\teffb$ show several trends- see Figure \ref{fig:TwoTempResults} and Appendix Table \ref{tbl_order_results} for the results of effective temperatures for all IGRINS orders.  \todo{Explain this more.}

Overplotting models with the observed spectra yielded insights on why some orders failed, and which orders can discriminate between starspots and ambient photospheric lines the best.  Some orders are dominated by spectral line outliers--deep metal lines that cannot be described by a coarse metallicity estimate alone.  These spectral line outlier orders also tend to bias estimates of effective temperature.  These metal lines can also be biased by Zeeman broadening, which would explain the heightened effect size with wavelength.

@@

Figures \ref{fig:Hband3x7} and \ref{fig:Kband3x7} in the Appendix show 42 of the $H-$ and $K-$band spectra on a log scale with a single random composite model spectrum overplotted.  The red and blue lines show the hot photosphere and cool spot spectra respectively.  We inspected the quality of the fits, generally finding good agreement with most spectral orders.  Many orders demonstrated spectral line outliers.  Most of the gross spectral features are captured by the composite model.  There are many examples of spectral lines that are present only in the hot photosphere model, or only in the cool photosphere model.  \todo{Insert new figure and text here.} For example, the OH 1.563 $\mu$m line noted by \citet{oneal01} shows a clear pattern of 3 lines in the data, with the central line exceeding the depth of the adjacent two lines.

Best fit values for the stellar parameters summarized are listed in Table \ref{tbl_adopted_props}.  Remarkably, the filling factor of starspots exceeds 50\% with a best fit value of $f_{\Omega}=75\pm 10 \% $.

@@

The two panels in Figure \ref{fig:lores_twopanel} show photospheric spectra scaled separately to the observed red-optical and near-IR spectra.  The single temperature spectra are unable to reproduce the observed emission.  The $3900$ K component fits the TiO bands reasonably well at $<7400$ \AA\ but fails to reproduce molecular bands at longer wavelengths.  The $3500$ K component hotter component suffers from the opposite problem, yielding molecular bands at short wavelengths that are too deep.  Only the best-fit photosphere from \S 3 fits both the optical and the near-IR bands well.

@@

\comm{GJH: needs revision.}
Photometry is obtained from the main sequence colors compiled by \citet{kenyon95}.  Colors from dwarf stars should have minimal contribution from spots; measurements of pre-main sequence colors will include some unknown contribution of spots and cannot be used for this analysis.  Synthetic spectra of $3000-4000$ K stars predict $B-V$ colors that are much bluer than observed and that become bluer at cooler temperatures, the opposite of what is observed.  As a consequence, synthetic photometry also cannot be used here.

@@

The rotational modulation leads to much smaller brightness changes at near-IR wavelengths than at optical wavelengths (see Table~\ref{tab:photrange}).  The smaller amplitude of near-IR brightness is also seen in a few spotted stars in the optical/IR monitoring of NGC 2264 \citep{cody14}, although those stars have a much smaller $V$ band amplitude than LkCa 4.  Figure~\ref{fig:colors} demonstrates that the $B-V$ color from \citet{grankin08} does not depend on $V$, while the $V-R_J$ becomes slightly redder when $V$ is smaller.  The predicted colors from rotational modulation do not match the observed $V-R_J$ colors or the correlation with $V$, either because of changes in spot properties since the \citet{grankin08} data were obtained or because a two-temperature model for the photosphere is overly simplistic.

@@

The TiO band depths in these spectra are measured from spectral indices listed in Table~\ref{tab:tio}.  The TiO indices are also combined into a single index by  calculating the number of standard deviations each point is from the median value of the index, and then averaging the standard deviation.

@@

Scaling our composite spectrum for LkCa 4 to the observed $K$-band photometry yields $R=2.2$ $R_\odot$ and $\log L/L_\odot=0.17$, assuming a distance of 131 pc for the parallax distance of the nearest Taurus member \citep{torres12}.   The spectrum of LkCa 4 is well reproduced by a cool component of $3000$ K covering 75\% of the surface and a $4200$ K component covering 25\% of the surface.  The effective temperature of these two components is 3611 K, as calculated from the average surface flux at the median $V$ band magnitude.  The apparent effective temperature in any single epoch would depend on the relative fractions of the hot and cool spot.  Decreasing the cool spot size to 60\% would increase the effective temperature to 3670 K, while increasing the spot size to 85\% would decrease the effective temperature to 3370 K.  {\bf GJH note:  holy crap!!!}.  Since the Hayashi tracks are nearly vertical at this temperature and age, the temperature uncertainty leads directly to an uncertainty in inferred mass.

@@

\subsection{Umbra versus penumbra}
Sunspots consist of cool penumbra enveloping much cooler and smaller umbra with temperature differences of X and Y respectively \todo{cite}.  We must be seeing the umbra of the star.

@@

\subsection{Starspots as a source of RV Jitter}
Starspots cause significant jitter in radial velocity surveys for young planets \citep[e.g.][]{donati14, robertson14}.  \todo{Potentially expand on this topic, but not critical.}

This observation has motivated precision radial velocity (RV) planet searches in the infrared, where the modulation of RV signals is lessened \citep{prato08}.
